---
title: "Aides Drug"
author: "Alfredo Acevedo"
date: "2/15/2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```


# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.
  
```{r}
drugies<-filter (triUmp, DrugUser == "yes")
```

  
# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->
  
```{r}
barchartGC(~infected + treatment,data=drugies)
barchartGC(~infected + treatment,data=drugies, type="percent")

```

  This shows that most of the people who take the drug have a better chance of preventing aids. Also shows that the placebo is not as effective as the drugs.
    
#### Plot barplots ggplot way

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
   
```{r}
table1 <- xtabs(~infected + treatment,data=drugies)
rowPerc(table1)
colPerc(table1)
```
   
This shows that the drugs help out almost three times the amount then the people taking the placebo. Meaning that the placebo is not good to help defend against the aids and people should take the drug instead.  
   
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
   
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  
  There are three results shown here and the two first one show a chi-square test and the other being a fisher's exact test. For the first chi-square test it shows that the p-value is extremely low, meanwhile the X-squared is about 28 which means that is to much critical value for the degree's of freedom which should be 1. There is a hugde difference in the data. 
  
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
  
As you can see from the graphic results and the infernal results. We can see that the people who had use the placebo had a higher chance of getting aids. Meanwhile people who used the drugs had almost tripled in the amount of them not having aids compared to the ones who took placebo and has aids. This means people shouldn't take the placebo.
  